---
title: "Task 4"
author: "Chris Gates"
date: "January 30, 2017"
output: html_document
---
Begin Task 4
Loading in ggplot2 and determining how many rows is in the diamonds data set in ggplot2
```{r}
library(ggplot2)
nrow(diamonds)
```

diamonds has 53,940 rows

```{r}
set.seed(1410)
dsmall <- diamonds[sample(nrow(diamonds), 100), ]
```

The set seed function will help to reuse the same set of random variables.
We need to declare set.seed before using the sample function in the next line.
dsmall is a random sampling of 100 rows of the 53,940 rows in the diamonds data set.

```{r}
plot1 <- ggplot(dsmall, aes(x = x, y = y, colour=z)) +
  geom_point() +
  facet_wrap(~cut)
print(plot1)
```

Make a series of scatter plots of x and y faceted by cut, z is added as an additional variable and is represented by color.

```{r}
plot2 <- ggplot(dsmall, aes(x = carat, y = price, color = cut)) +
geom_point() +
geom_smooth(method=lm,
            se=FALSE)
print(plot2)
```

A scatterplot of price vs carat, colored by cut and smoothed using the "lm" method (linear regression) and without standard error bars

```{r}
plot3 <- ggplot(dsmall, aes(carat, color = clarity)) +
  geom_density() +
  facet_wrap(~clarity)
print(plot3)
```

Density plot of carat, faceted and colored by clarity.

```{r}
plot4 <- ggplot(dsmall, aes(x = cut, y = price)) +
  geom_boxplot()
print(plot4)
```

A boxplot of price as a function of cut.

```{r}
plot5 <- ggplot(dsmall, aes(x = x, y = y, color)) +
  geom_point(colour = "red") +
  geom_smooth(method = lm, color = "blue", linetype = 2, se=FALSE) +
  coord_fixed() +
  xlab("x, in mm") +
  ylab("y, in mm")
print(plot5)
```

Scatterplot of y versus x, smoothed using linear regression line "lm" method.
Regression line is colored blue and made to be a dashed line.
Data points are colored red.
X and Y axis titles were changed to represent scale.

```{r}
cadata <- readxl::read_excel("C:/Users/Chris/Desktop/WInter/Wetting Angle Data/CM Data/CM-S1-1A_2017-01-08_05-21-56-PM.xlsx", sheet=1, col_names=TRUE)
cadata <- data.frame(cadata)
CAM <- data.frame(CA=c(cadata$Var.6), Time=c(cadata$Var.10))
CAM <- CAM[!is.na(as.numeric(as.character(CAM$CA))), ]
CAM <- data.frame(CAM)
CAM$CA=as.numeric(levels(CAM$CA))[CAM$CA]
CAM$Time=as.numeric(levels(CAM$Time))[CAM$Time]

CAplot <- ggplot(CAM, aes(Time, CA)) +
  geom_point() +
  scale_y_continuous(limits=c(50, 75)) +
  scale_x_continuous(limits=c(0, 100))

CAplot <- CAplot + geom_smooth(method=lm)
print(CAplot)
```

